From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-02 WOLFHPC15/article/ms_msc.tex
Forgotten files

\subsubsection*{Computation Description}
Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}. However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language. The exact code of the computation as to be written into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}). Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d$ in the formal definition), an identifier for the computation kernel, and the set of identifiers for the data read by the computation, with parenthesis. For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil). If the computation is local no brackets appear ($R_l$ in the formal definition). In the example of Figure~\ref{fig:mslex}, nine computations are described. For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c_0} which read data \texttt{a} without neighborhood shape.

@@

\subsubsection*{The Minimal Series-Parallel Graph: $\Gamma_{msp}$}
Once a dependency graph is built, many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}. In this work a static scheduling of the dependency graph is built. To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}. As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the dag. Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fg:over}.

>>>>

\subsubsection*{Computation Description}
Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}. However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language. The exact code of the computation as to be written into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}). Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d$ in the formal definition), an identifier for the computation kernel, and the set of identifiers for the data read by the computation, with parenthesis. For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil). If the computation is local no brackets appear ($R_l$ in the formal definition). In the example of Figure~\ref{fig:mslex}, nine computations are described. For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

@@

\subsubsection*{The Minimal Series-Parallel Graph: $\Gamma_{msp}$}
Once a dependency graph is built, many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}. In this work a static scheduling of the dependency graph is built. To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}. As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the dag. Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.

### 2015-08-02 WOLFHPC15/article/mscac.tex
Forgotten files

The solution proposed in this paper is based on a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a component-based back-end code.

@@

As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases. The first phase \emph{parses} the MSL description into an intermediate representation (IR). The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism of the overall numerical simulation. Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.

@@

%In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks. From a dependency graph, different strategies can be used to schedule tasks, in other words to control their execution in a valid order. The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.

The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation. We could have proposed a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime. Component models, where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes. Those properties, on the other hand, are recurrent stated problems of parallel applications.

>>>>

The solution proposed in this paper is based on \textbf{a)} a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and \textbf{b)} a compiler which successively parses the language, applies transformations for automatic parallelization, and finally dumps the representation to a component-based code.

@@

As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases. The first phase \emph{parses} the MSL description into an intermediate representation (IR). The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism in the overall numerical simulation.
Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.

@@

%In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a graph of tasks with dependencies.
Different strategies can be used to schedule the tasks of this graph; in other words to plan their execution in a valid order. The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.

The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
Various models could be used as back-end, such as OpenMP~\cite{660313} for example, however this paper adopts a novel approach where the dump targets a component-based runtime.
This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
%where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for

\todo[inline]{JB: je déplacerais les 2 derniers paragraphes et la figure dans la sec.~\ref{sect:component}}
